Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 22:14

spp <- rename(spp, "Trend [%/year]" = slope       )
spp <- rename(spp, "Sky condition"  = DATA        )
spp <- rename(spp, "Pearson cor."   = cor.estimate)

@@

Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions, and for each season of the year, using the corresponding mean monthly anomalies
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

>>>>

spp <- rename(spp, "Trend [%/year]" = slope       )
spp <- rename(spp, "Trend p-value"  = slope.p     )
spp <- rename(spp, "Sky condition"  = DATA        )
spp <- rename(spp, "Pearson cor."   = cor.estimate)

@@

Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions and for each season of the year, using the corresponding mean monthly anomalies
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant at the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 22:14

Similarly to the long term trends from daily means of SDR discussed
above, we have calculated the trend for the three sky conditions, and
for each season of the year, using the corresponding mean monthly
anomalies (Figure~\ref{fig:seasonalALL} and
Table~\ref{tab:trendseasontable}). For all-sky conditions the trend in
SDR in winter is the largest (\(0.7\,\%/y\)), followed by the trend in
autumn (\(0.42\,\%/y\), a value close to the long term trend) both
statistically significant above the \(99\,\%\) confidence level. In
spring and summer, the trends are much smaller and of lesser statistical
significance. These seasonal differences indicate a possible relation of
the trends in SDR to trends of clouds during winter and autumn. For
clear-skies, the trend in winter is \(0.4\,\%/y\), larger than for
all-skies (\(0.7\,\%/y\)), which is another indication of a decreasing
trend in cloud optical thickness. Moreover, the trends under clear- and
cloudy-sky conditions are almost complementary to each other,
particularly for winter and autumn, where the signal is stronger. During
spring and summer the statistical significance is very low and the
actual trend too small for a meaningful comparison.

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
\toprule
Trend [\%/year] & slope.p & Pearson cor. & Sky condition & Season\\
\midrule
\cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.002770} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter}\\
\cmidrule{1-5}
0.1070 & 0.371000 & 0.1660 & All skies & Spring\\
\cmidrule{1-5}
\cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.175000} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer}\\
\cmidrule{1-5}
0.4230 & 0.009390 & 0.4660 & All skies & Autumn\\
\cmidrule{1-5}
\cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.000515} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter}\\
\cmidrule{1-5}
0.0591 & 0.497000 & 0.1270 & Clear skies & Spring\\
\cmidrule{1-5}
\cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer}\\
\cmidrule{1-5}
0.0539 & 0.366000 & 0.1710 & Clear skies & Autumn\\
\cmidrule{1-5}
\cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.000665} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter}\\
\cmidrule{1-5}
0.0621 & 0.593000 & 0.0998 & Cloudy skies & Spring\\
\cmidrule{1-5}
\cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.560000} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer}\\
\cmidrule{1-5}
0.1910 & 0.384000 & 0.1650 & Cloudy skies & Autumn\\
\bottomrule
\end{tabu}
\end{table}

>>>>

Similarly to the long term trends from daily means of SDR discussed
above, we have calculated the trend for the three sky conditions and for
each season of the year, using the corresponding mean monthly anomalies
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
all-sky conditions the trend in SDR in winter is the largest
(\(0.7\,\%/y\)), followed by the trend in autumn (\(0.42\,\%/y\), a
value close to the long term trend) both statistically significant at
the \(99\,\%\) confidence level. In spring and summer, the trends are
much smaller and of lesser statistical significance. These seasonal
differences indicate a possible relation of the trends in SDR to trends
of clouds during winter and autumn. For clear-skies, the trend in winter
is \(0.4\,\%/y\), larger than for all-skies (\(0.7\,\%/y\)), which is
another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost
complementary to each other, particularly for winter and autumn, where
the signal is stronger. During spring and summer the statistical
significance is very low and the actual trend too small for a meaningful
comparison.

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky condition & Season & Trend [\%/year] & Pearson cor. & Trend p-value\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{0.002770}\\

 & Spring & 0.1070 & 0.1660 & 0.371000\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{0.175000}\\

\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4230 & 0.4660 & 0.009390\\
\cmidrule{1-5}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{0.000515}\\

 & Spring & 0.0591 & 0.1270 & 0.497000\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{0.106000}\\

\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.0539 & 0.1710 & 0.366000\\
\cmidrule{1-5}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{0.000665}\\

 & Spring & 0.0621 & 0.0998 & 0.593000\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{0.560000}\\

\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.1910 & 0.1650 & 0.384000\\
\bottomrule
\end{tabu}
\end{table}
